Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-02-01 text/src/article.tex
Revise scaffold file section of results

The genome scaffold is written in YAML format. YAML is a data format designed
to be both machine readable and easy to manually edit. An example of this
format illustrating all the features of scaffolder is shown in Figure 1. \pb

@@

The first line of the file begins with three dashes. This is required to
signify the start of a YAML document. Lines 2 begins with a dash character `-'
which denotes the start of an entry in the sequence scaffold, in this case the
first entry. Line 3 is indented by two spaces which therefore means this line
contains an entry type. The ``sequence'' tag is used indicate a sequence is
inserted at this region in the scaffold. The following line contain is
indented by two spaces and indicates that this an attribute of the sequence
entry. In this case the ``source'' keyword is used to indicate used which
sequence should be inserted into the scaffold. This sequence is identified by
the first word of the fasta header of the selected sequence. This indicates
a simple sequence region added to the scaffold. \pb

@@

The following line has no indentation and begins with dash indicating the next
entry in the scaffold. The following entry is identified by the ``unresolved''
tag. This identifies a region of unknown nucleotide sequence but is included
because the approximate size of the region is known. The following line defines
that the size of this unknown region in the genome scaffold should be 20
nucleotides. This therefore results in the region being filled with 20 `N'
characters in draft sequence. \pb

@@

The following entry is another sequence using the sequence from 'sequence2' in
the scaffold file. This sequence entry has three further attributes in addition
to defining the source sequence. The first two are start and stop. These are
used to trim the final sequence inserted into the scaffold. In this case the
first two nucleotides of the sequence are removed so that that sequence starts
at position 3 of the original raw sequence. Similarly the sequence after
position is trimmed away after position 100. These sequence is also reverse
complemented using the ``reverse'' tag. Reverse complementation is performed
after the sequence has been trimmed to the specified coordinates. \pb

The final attribute for this sequence entry defines inserts to be added to the
sequence. These are additional regions of sequence that can be added to
a sequence entry to alter the existing sequence such to fill gaps or correct
sequencing errors. Inserts are defined using the ``inserts'' tag. Each insert
is then listed similarly to the entries in the scaffold: beginning with a dash
on one line then indented by an additional two spaces on the following lines.
\pb

Inserts are listed similarly to sequence entries where reverse, start and stop
attributes can be used to trim and reverse the sequence. The `source' tag is
used to identify which fasta sequence entry should be used as the insert.  The
insert has two additional tags ``open'' and ``close'' which determine where the
insert is added to the hosting sequence. These coordinates are inclusive
replacing the coordinate nucleotides. The open and close positions need not be
the same size as the insert size. Any difference in size to the replaced
nucleotides will result in the length on the encoding sequence being adjusted.
\pb

The next insert entry shows an insert with only the open position specified.
This is to illustrate that only the one of the `open' or `close' keywords is
sufficient to add an insert. If only one of these words is used then the
opposing insert position is calculated from the length on the insert sequence
in the fasta file. This feature is designed for using inserts to close gaps
where the length of the insert may not completely fill the gap but can still be
used to fill a portion of the gap sequence. \pb

>>>>

The scaffold file is written using `YAML', a data format designed to be both
machine readable and manually editable. An example of a scaffold file in YAML
format is shown in Figure 1. This example scaffold file illustrates the
available features in scaffolder. The basic layout of the scaffold file is as
a list - a series of entries where each entry defines a region of sequence to
be generated for the final super sequence. In the final super sequence each of
the entries are joined together to generate a single continuous sequences. \pb

@@

The first line of the scaffold file begins with three dashes. This is
a requirement to indicate the start of a YAML formatted document. Line 2 begins
with a dash character `-' denoting the first entry in the sequence scaffold.
Line 3 is indented by two spaces. Whitespace is used to delimite attributes in
the file. Lines with an incremented level of white space are considered linked
to the first preceeding non-indented line. \pb

The ``sequence'' tag used on line 3 indicates that this region is in the
scaffold will contain nucleotide sequence from the fasta file. The following
line indented by two spaces and indicating an attribute of the sequence entry
defined on line 3. In this case the ``source'' keyword identifies the name of
the sequence to be inserted into the scaffold. As described above the first
space-delimited word of the fasta header should be used to identify the
sequence. Taken together lines 2-4 describe the first entry in the scaffold is
a simple sequence which the identifier `sequence'. \pb

@@

Line 5 has no whitespace indentation and begins with a dash and therefore
describes the next entry in the scaffold. This entry is identified by the
``unresolved'' tag, and specifies a region of unknown nucleotides but known
length. The following line specifies the size of this unknown region. Lines 5-6
therefore insert a region of 20 `N' characters in the super sequence directly
following the region defined by the first entry on lines 2-5. \pb

@@

The last entry adds the sequence 'sequence2' in the scaffold file. This entry
demonstrates three further attributes for a sequence entry.The `start' and
`stop' tags are used to trim the sequence used in the scaffold to these
coordinates inclusively. The ``reverse'' tag is used to reverse complement this
region in the final super sequence. \pb

This region further includes the ``inserts'' tag to update the inserted
sequence with additional regions of sequence. These inserts are also added as
a YAML list with each insert listed with a dash on one line followed by the
remainder of the insert errata. \pb

Lines 16-19 illustrate an insert using similar attributes to that of a sequence
entry. The reverse, start and stop tags trim the inserted sequence and the
`source' tag identifies the corresponding fasta sequence. The ``open'' and
``close'' tags on lines 20-21 determine where the insert is added to the
enclosing sequence.  These coordinates inclusively replace the region within
these coordinates with the specified fasta sequence. The open and close
positions need not be the same size as the inserted sequence. \pb

The next insert on lines 23-24 specifies only the open position. This
illustrate that only the one of the `open' or `close' tags is required.  When
only open or close tag is used then the opposing position is calculated from
the length on the insert fasta sequence. This allows inserts to partially fill
gap regions in sequences. \pb
